---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align="center",
  dpi = 300
)
```

# pedlikCompare

The purpose of **pedlikCompare** is to compare pedigree likelihoods computed by different programs. Such comparisons are invaluable for software developers, giving numerical validation of the implemented algorithms. Furthermore, comparing runtimes may identify and document strengths and weaknesses of each programs, potentially pinpointing computational bottlenecks that may be improved.

Several R packages are able to calculate pedigree likelihoods, including **pedprobr** (part of the [ped suite](https://magnusdv.github.io/pedsuite/)), **paramlink** (precursor to the ped suite), **Familias**, and **ElstonStewart**. Outside of R, a widely used program is [**MERLIN**](https://csg.sph.umich.edu/abecasis/Merlin/). The **pedlikCompare** package makes it easy and fun to compare all of these programs, both in terms of numeric accuracy and runtime. (For MERLIN to be included it must be installed on your computer.)

**pedlikCompare** imports **pedtools** for creating pedigrees and markers. Conversion to all other formats are done automatically when needed.

## Example

Consider the pedigree shown below, where a child and its parents have been genotyped with a SNP marker. We set this up with **pedtools** as follows:

```{r}
library(pedtools)

x = nuclearPed() |> 
  addMarker(geno = c("1/2", "1/2", "2/2"), afreq = c(`1` = 0.5, `2` = 0.5))
```

Plot the pedigree to check that everything is ok.
```{r example, fig.width = 2.5, fig.height = 2.5, out.width = "35%"}
plot(x, marker = 1)
```

Now let **pedlikCompare** perform its magic! The crucial function is `compare()`.
```{r}
library(pedlikCompare)
result = compare(x)
```

As indicated in the output, all programs agreed in this case. The `result` object contains more details:
```{r}
result
```

In order to compare the likelihoods, `compare()` calls the function `all_agree()` which deals with rounding and other mundane issues. If you happen to know the exact likelihood, this can be supplied in the optional `answer` argument. In our example it is `1/16`, so the command becomes:
```{r}
all_agree(result, answer = 1/16)
```

